Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: Added some comments and suggested changes.

\title*{Challenges in Developing Great Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al}
\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and R. Jagadeeswaran Aleksei G. Sorokin}
\institute{
Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
\and
Yuhan Ding \at
 Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{yding2@hawk.iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and
R. Jagadeeswaran \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@hawk.iit.edu}
\and
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

@@

\abstract{
Quasi-Monte Carlo (QMC) methods have developed over several decades. With the explosion in computational science, there is a need for great software that implements QMC algorithms. We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
}

@@

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries could provide more flexible LD offerings.  Some large environments, such as  Dakota \cite{DakotaUsersManual} and SAS, have quite limited LD generators and should expand their offerings.  Developers of smaller libraries with more QMC features, like BRODA and QMCPy, should demonstrate how their libraries can connect well with other software environments.  An example of QMCPy connecting with other software is given in the next subsection.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five. These plots may be reproduced using the Jupyter notebook \SCComment{Add url}.}
    \label{fig:beam}
\end{figure}

>>>>

\title*{Challenges in Developing Great Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al.}
\author{Sou-Cheng T.~Choi \and Yuhan Ding \and Fred J. Hickernell \and Jagadeeswaran  Rathinavel \and Aleksei G. Sorokin}
\institute{
Sou-Cheng T.~Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@iit.edu}
\and
Yuhan Ding \at
 Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{yding2@hawk.iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and
Jagadeeswaran  Rathinavel \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@hawk.iit.edu}
\and
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

@@

\abstract{
Quasi-Monte Carlo (QMC) methods have developed over several decades. With the explosion in computational science, there is a need for great software that implements QMC algorithms. We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software. \SCComment{We demonstrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.}
}

@@

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible means?} LD offerings.  Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings. \SCComment{Other established software such as SAS expand offerings of trustworthy and efficient analytics through continuous internal implementation as well as flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}. Such latter strategy represents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.  An example of QMCPy connecting with other software is given in the next subsection.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up of factor four to five. These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url. (2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)? Are there numerical values on (left and right) y-axis? (3) The font size of the axis labels for the lower graphs could be made larger more like the top graph. (4) I guess the blue colors are covered by the orange colors in the bottom right graph. Maybe we can use blue crosses and orange broken lines instead.}}
    \label{fig:beam}
\end{figure}
